\section{The Governing Law in the Federal Courts}

\begin{enumerate}
    \item Holmes: ``The common law is not a brooding omnipresence in the sky, 
    but the articulate voice of some sovereign or quasi sovereign that can be 
    identified~.~.~.~''\footnote{\emph{Southern Pac. Co. v. Jensen} 244 U.S. 205, 
    222 (1917).}
    \item \textbf{``Rules of Decision Act''} determines which law applies in 
    federal courts: ``The laws of the several states, except where the 
    Constitution or treaties of the United States or Acts of Congress 
    otherwise require or provide, shall be regarded as rules of decision in 
    civil actions in the courts of the United States, in cases where they 
    apply.\footnote{28 U.S.C. \S\ 1652, originally codified in civil \S\ 34 of 
    the Judicial Act of 1789. See Casebook p. 489.}
    \item The \textbf{Rules Enabling Act} (28 U.S.C. \S\ 2071--2077) gives the 
    federal court system the power to adopt federal rules. The act stipulates 
    that procedural rules \textbf{``shall not abridge, enlarge or modify any 
    substantive right.''} \S\ 2072.
    \item The Supreme Court's interpretation of the Rules of Decision Act 
    differed sharply betwen \emph{Swift} (the old rule) and \emph{Erie} (the 
    new rule). Previously, federal courts applied \textbf{general federal 
    common law}. After \emph{Erie}, federal courts apply the law that the 
    highest court in the state would apply.\footnote{This is often the law of 
    the state in which the court sits. Often, though, other factors require 
    application of another state's law, such as in \emph{Erie}, where the New 
    York federal court applied Pennsylvania state law.}
    \item If disregarding state law would significantly affect the outcome of 
    the case, the \textbf{outcome determinative test} requires the court to 
    apply state law. \emph{York}. However, the Court scaled back the test in 
    \emph{Byrd}, where it held that strong federal interests can weigh in 
    favor of applying state law.
    \item In \emph{Hanna}, the Court held that courts should apply federal 
    \emph{procedural} rules if they conflict with state rules. (The 
    \emph{Erie} doctrine still applies to \emph{subtantive} law.) The Court 
    developed a two-part test to determine whether to apply state or federal 
    procedural law. See below for the full test.
    \item A federal procedural rule displaces a state rule \emph{only if} 
    there is an unavoidable conflict between the two. \emph{Walker}.
    \item Federal common law still applies in some limited circumstances. 
    \emph{Clearfield}.
\end{enumerate}

\subsection{General Law over State Law: \emph{Swift v. Tyson}}

Where state law deviates from general law, federal courts should follow 
general federal common law, \emph{not} state common law.

\begin{enumerate}
    \item Swift owned a bill of exchange that Tyson originally made to two 
    other men, who endorsed it over to Swift. Tyson refused to pay because of 
    a breach of the original contract for which the bill of exchange was 
    originally issued.
    \item Swift sued Tyson in federal court for payment. Under 
    ``local`` law (New York state law), which the state court followed, 
    Tyson's defense was valid. Under ``general'' law, which federal courts 
    followed, the defense was not valid against Swift.
    \item \S\ 34 of the Judiciary Act of 1789 required the application of ``the 
    laws of the several states\ldots in cases where they apply'' unless 
    federal law otherwise required (roughly equivalent to 28 U.S.C. \S 1652).
    \item In a unanimous opinion, Justice Story held that the federal court 
    ``should follow the general law rather than a state's local law in cases 
    where the state law deviated from the general law.''\footnote{Casebook p. 
    491.}
    \item After the Civil War, the economic interests of the states began to 
    diverge (north: finance, manufacturing; south: agriculture, labor). The 
    Supreme Court expanded general law to include torts, so industrial 
    accidents were increasingly litigated in federal courts. Federal courts 
    also grew increasingly sympathetic to creditors and employers. Those who 
    favored the results of state courts became enemies of \emph{Swift}.
\end{enumerate}

\subsection{No Federal General Common Law: \emph{Erie R.R. Co. v. Tompkins}}

``Except in matters governed by the Federal Constitution or by acts of 
Congress, the law to be applied in any case is the law of the state~.~.~.~.  
\textbf{There is no federal general common law}.''\footnote{Casebook p. 497.}

\begin{enumerate}
    \item A passing train injured Tompkins was while he was walking on a 
    footpath along a railroad track in Pennsylvania. He brought suit against 
    the Erie Railroad Co. in federal court in New York.\footnote{Prof. Bradt, 
    10/2/12: ``One note clarifying something at the end of our discussion 
    today. I believe that Mr. Mornin wanted to know why Harry Tompkins wanted 
    to filed his case in federal court in New York as opposed to federal court 
    in Pennsylvania.

    He could have filed in federal court in either state (based on diversity), 
    there was likely personal jurisdiction over the Erie in Pennsylvania (even 
    under 1930s-era personal-jurisdiction doctrine, the Erie would have been 
    considered to be ``doing business'' in Pennsylvania, at least when it came 
    to an accident in Pennsylvania), and venue would have been fine, too (in 
    those days, venue could be based on the residence of the plaintiff).

    The decision to file in federal vs. state court is clear: Tompkins wanted 
    the benefit of federal general common law, as opposed to PA state law, 
    which was terrible for him. But why NY federal court as opposed to PA 
    federal court?

    Well, the answer lies in the lack of uniformity of general federal
    common law. Not only did \emph{Swift} not achieve uniformity between 
    federal and state courts on issues of tort law, it also did not achieve 
    uniformity \emph{among} federal courts. A federal court sitting in 
    Pennsylvania (governed by the law of the Third Circuit) would have applied 
    more defense-friendly law than the federal court sitting in New York 
    (governed by the Second Circuit). It's possible, in fact, that the PA 
    federal district court would have deferred to PA law in Tompkins's case 
    because the Third Circuit's view of \emph{Swift} was a little less 
    expansive than the Second Circuit's. So even though general federal 
    common law would have applied in federal court sitting in PA and NY, the 
    NY federal court would have been a friendlier forum.''}
    \item Under Pennsylvania state law, Tompkins would have been considered a 
    trespasser and therefore not entitled to recover damages. Under general 
    law, the railroad might have been held negligent.
    \item The legal circumstances were unusual. At the time, ``general law'' 
    usually benefited corporations, while ``local law'' usually favored 
    individuals. In this case, however, the plaintiff argued for the 
    application of general law.
    \item The trial court and appellate court, following \emph{Swift}, found 
    that since no state statute governed the issue at hand, general law should 
    control. They found for the plaintiff. The issue before the Supreme Court 
    was whether the district court was free to disregard Pennsylvania common 
    law.
    \item Justice Brandeis:
    \begin{enumerate}
        \item The \emph{Swift} court had misinterpreted the intentions of the 
        authors of the Judiciary Act of 1789: ``~.~.~.~the construction given to 
        it by the court was erroneous; and that the purpose of the section was 
        merely to make certain that, in all matters except those in which some 
        federal law is controlling, the federal courts exercising jurisdiction 
        in diversity of citizenship cases would apply as their rules of 
        decision the law of the state, unwritten as well as 
        written.''\footnote{Casebook p. 495.} (The notes point out that the 
        historian on whom the opinion relies, Charles Warren, might have 
        mistakenly interpreted the Rules of Decision Act as requiring federal 
        courts to follow state common law even in areas where federal common 
        law applied.)
        \item \emph{Swift} caused significant ``injustice and 
        confusion''\footnote{Casebook p. 501}---e.g., companies 
        reincorporating in other states in order to establish diversity 
        jurisdiction to have their cases tried in federal court (\emph{Black 
        \& White Taxicab}). ``\emph{Swift v. Tyson} introduced grave 
        discrimination by noncitizens against citizens.''\footnote{Casebook p. 
        496.}
        \item The federal government did not have the power to legislate rules 
        of tort or contract law.\footnote{This quickly became untrue as the 
        Court expanded the federal government's power to regulate these areas 
        under the Commerce Clause.} Federal courts also do not have the power 
        to create rules in these areas.
        \item Reversed and remanded to be decided on the basis of Pennsylvania 
        state law.
        \item The \emph{Swift} rule is overturned.
    \end{enumerate}
    \item ``Except in matters governed by the Federal Constitution or by acts 
    of Congress, the law to be applied in any case is the law of the state~.~.~.~.  
    There is no federal general common law.''\footnote{Casebook p. 497.} 
    Judges often relied on ``general law'' as a way of ignoring state laws 
    that conflicted with their views.
    \item Justice Reed, concurring:
    \begin{enumerate}
        \item It is enough to broaden the \emph{Swift} framework to hold that 
        ``the laws'' in the Rules of Decison Act include state common law, 
        rather than declare the entire \emph{Swift framework} to be 
        unconstitutional.
        \item It's ``questionable'' to say that Congress has no power to 
        declare which substantive laws control in federal courts---moreso 
        because ``[t]he line between procedural and substantive law is 
        hazy.''\footnote{Casebook p. 499.}
    \end{enumerate}
\end{enumerate}

\subsection{``Outcome Determinative'' Test: \emph{Guaranty Trust v. York}}

The \textbf{``outcome determinative'' test}: would it significantly 
affect the outcome of the litigation for a federal court to disregard state 
law? If so, \emph{Erie} holds that the court should follow state law.

\begin{enumerate}
    \item Background:
    \begin{enumerate}
        \item \textbf{Substance vs. procedure}: did \emph{Erie} and the Rules 
        of Decision Act apply to both procedural and substantive law?
        \item Before the FRCP were enacted in 1938, courts were divided into 
        courts of law (in which cases were triable by jury) and courts of 
        equity (where there was no right to a jury).
        \item The \textbf{Conformity Act of 1872} required that in cases 
        \emph{at law}, federal courts must conform to the procedural rules of 
        the states in which they were located. Thus, there were procedural 
        differences between federal courts in different states, but few 
        differences between federal and state courts in the same state. In 
        cases \emph{in equity}, federal and state courts followed different 
        rules. Federal courts developed their own system of procedural rules 
        for suits brought in equity.\footnote{Casebook p.  504.}
        \item The \textbf{Rules Enabling Act of 1934} authorized the Supreme 
        Court (with congressional approval) to develop, with the consent of 
        Congress, a national system of procedural rules for federal civil 
        cases. So while \emph{Erie} required federal courts to follow state 
        rules in substantive law, federal courts developed independent 
        procedural rules (codified in the FRCP in 1938).
    \end{enumerate}
    \item York brought suit in 1942 equity in New York federal court for fraud 
    that occurred in 1931. The defendants argued that the New York statute of 
    limitations applied to cases both at law and in equity. The plaintiff 
    argued that the federal rule of laches, which typically applied in equity 
    cases, should apply in this case.
    \item The trial court agreed with the plaintiff and applied the New York 
    statute of limitations. The appellate court reversed, holding that the 
    federal laches doctrine should have applied, and granted summary judgment 
    to the defendants.
    \item Justice Frankfurter:
    \begin{enumerate}
        \item There is not a clear distinction between ``substantive'' and 
        ``procedural'' rights.
        \item This case dealt with a state-created right. When a federal court 
        adjudicates a state-created right solely on the basis of jurisdiction, 
        it becomes ``in effect, only another court of the State.''
        \item Does the statute of limitations affect ``merely the manner and 
        the means'' of the right to recover, or is it ``a matter of 
        substance'' that affects the result of the litigation?
        \item \emph{Erie} did not intend to ``formulate scientific legal 
        terminology'' around the terms ``substantive'' and ``procedural.'' It 
        intended to ensure that outcomes of diversity cases in federal court 
        would be similar to outcomes in state courts. If outcomes were 
        different, we're back to the discriminatory days of \emph{Swift}.
        \item The \emph{Erie} doctrine does not distinguish between cases at 
        law and cases in equity.
        \item ``The source of substantive rights enforced by a federal court 
        under diversity jurisdiction, it cannot be said too often, is the law 
        of the States. Whenever that law is authoritatively declared by a 
        State, whether its voice be the legislature or its highest court, such 
        law ought to govern in litigation founded on that law, whether the 
        forum of application is a State or a federal court and whether the 
        remedies be sought at law or may be had in equity.''\footnote{Casebook 
        p. 507.}
        \item Reversed and remanded for hearings consistent with the New York 
        statute of limitations.
    \end{enumerate}
    \item Justice Rutledge, dissenting:
    \begin{enumerate}
        \item The distinction between ``substantive'' and ``procedural'' law 
        is arbitrary but important.
        \item Forum states are free to apply their own statutes of 
        limitations, which may be different from those of the state that 
        originally created the substantive right.
    \end{enumerate}
\end{enumerate}

\subsection{Really Regulating Procedure: \emph{Sibbach v. Wilson \& 
Co.}}

According to the Rules Enabling Act, the federal rules ``shall neither 
abridge, enlarge, nor modify the substantive rights of any litigant.'' A rule 
is procedural if it ``really regulates procedure.''

\begin{enumerate}
    \item A federal court required the plaintiff to submit to a mandatory medical 
    examination. The examination would not have been mandatory under Illinois 
    state law. She refused the exam on the grounds that the Rules Enabling Act 
    forbade rules that abridge litigants' substantive rights. The court held 
    ``[t]he test must be whether a rule \textbf{really regulates 
    procedure}.''\footnote{Casebook p. 509.}
    \item Justice Frankfurter (and three others) dissented, arguing that the 
    examination rule constituted ``invasion of the person'' and was therefore 
    significantly different from other procedural rules.
\end{enumerate}

\subsection{``Byrd Balancing'': \emph{Byrd v. Blue Ridge Rural Elec. Coop.}}

\textbf{Byrd balancing}: even if a state rule is outcome determinative, 
countervailing federal considerations can tip the balance in favor of applying 
federal rules.

\begin{enumerate}
    \item Byrd sued the defendant in North Carolina district court for 
    negligence for an injury he sustained while connecting power lines.
    \item The defendant argued that the plaintiff was a ``statutory employee'' 
    under the South Carolina Workmen's Compensation Act when the injury 
    occurred, which would mean that the plaintiff was barred from suing and 
    was obliged to accept statutory compensation benefits.
    \item The district court found for Byrd. The Fourth Circuit reversed.
    \item The questions before the Supreme Court were (1) whether the 
    appellate court erred in directing judgment for Blue Ridge without giving 
    Byrd an opportunity to introduce further evidence, and (2) whether Byrd is 
    entitled to a jury to determine factual issues.
    \item Justice Brennan:
    \begin{enumerate}
        \item Blue Ridge argued that a judge, not a jury, should decide the 
        question of immunity. In \emph{Adams v. Davidson-Paxon Co.}, the South 
        Carolina Supreme court held that a judge, not a jury, should determine 
        the question of whether the plaintiff was a statutory employee (and 
        therefore whether the employer is immune from paying damages). The 
        defendant's argument contends that the federal court should follow 
        this state precedent.
        \item \emph{Erie} held that in diversity cases, federal courts 
        must respect state-created rights by state courts. Here, the Supreme 
        Court found that the decision in \emph{Adams} to send the immunity 
        question to a judge was a ``practical consideration''---``merely a 
        form and mode''---``and not a rule intended to be bound up with the 
        definition of the rights and obligations of the parties.''
        \item Mere ``form and mode'' rules can be important if they 
        bear substantially on the outcome of the litigation (cf. 
        \emph{York}). It may be in this case that the question of whether 
        immunity should be decided by a judge or jury would bear substantially 
        on the outcome. Here, however, there is a general federal policy of 
        sending factual questions to the jury. Even though it may affect the 
        outcome, the federal rule should outweigh the state rule because a 
        federal procedural rule is ``not in any sense a local matter.''
        \item Third, it is not at all clear that sending the immunity question 
        to a jury would result in a different outcome.
        \item The federal rule applies. Reversed and remanded.
    \end{enumerate}
\end{enumerate}

\subsection{State vs. Federal Procedural Rules: \emph{Hanna v. Plumer}}

The ``\emph{Hanna} holding'' and ``\emph{Hanna} dictum'' tests determine when 
the court should follow state or federal procedural rules.

\begin{enumerate}
    \item Hanna was involved in an auto accident in South Carolina. He sued 
    the executor of the other driver's estate, Plumer, in Massachusetts 
    district court.
    \item In accord with the Federal Rules, Hanna served process by leaving 
    copies of the summons with the defendant's wife at their residence. Plumer 
    argued that the conflicting Massachusetts statute requiring service in 
    person should apply. If the Massachusetts standard applied, the statute of 
    limitations would have run.
    \item The district court granted summary judgment for Plumer and the First 
    Circuit affirmed.
    \item Justice Warren developed a two-prong test for federal procedural 
    rules:\footnote{The Supreme Court offered a slightly different framing of 
    the \emph{Hanna} test in \emph{Burke v. Smith}:
    \enquote{~.~.~.~the Supreme Court developed a two-part test in \emph{Hanna v.  
    Plumer}~.~.~.~. Under the \emph{Hanna} test, \enquote{when the federal law 
    sought to be applied is a congressional statute or Federal Rule of Civil 
    Procedure, the district court must first decide whether the statute is 
    `sufficiently broad to control the issue before the court.'}~.~.~.~  
    If the federal procedural rule is \enquote{sufficiently broad to control 
    the issue} and conflicts with the state law, the federal procedural rule 
    applies instead of the state law~.~.~.~. If the federal rule does not directly 
    conflict with the state law, then the second prong of the \emph{Hanna} 
    test requires the district court to evaluate \enquote{whether failure to 
    apply the state law would lead to different outcomes in state and federal 
    court and result in inequitable administration of the laws or forum 
    shopping.}}}
    \begin{enumerate}
        \item What is the source of the procedural rule?
        \begin{enumerate}
            \item Congressional rule (``\emph{Hanna} holding''):
            \begin{itemize}
                \item Is the rule \textbf{pertinent}? I.e., is there an 
                unavoidable conflict with a state procedural rule?
                \item Is the rule \textbf{valid}?\footnote{No FRCP has ever 
                been found to be invalid, thought the Court has interpreted 
                them narrowly to preserve their validity.}
                \begin{itemize}
                    \item Is it \textbf{constitutional}? I.e., is it 
                    rationally capable of classification as procedural?
                    \item Is it consistent with the \textbf{Rules Enabling 
                    Act}? I.e., does it ``really regulate procedure'' 
                    (\emph{Sibbach}) without abridging substantive rights?
                \end{itemize}
                \item $\rightarrow$ If yes, apply the congressional rule. 
                Otherwise not.
            \end{itemize}
        \end{enumerate}
        \begin{enumerate}
            \item Judge-made rule (``\emph{Hanna} dictum''):
            \begin{itemize}
                \item Is the rule \textbf{outcome determinative} 
                (\emph{Guaranty})?
                \item Does the rule implicate both of the \textbf{twin aims of 
                \emph{Erie}}?
                \begin{itemize}
                    \item Discourage \textbf{forum shopping}.
                    \item Avoid \textbf{inequitable administration of the 
                    law}. Can you explain to your client why being in federal 
                    vs. state court makes a difference?
                \end{itemize}
                \item $\rightarrow$ If yes, apply the judge-made rule. 
                Otherwise.
            \end{itemize}
        \end{enumerate}
    \end{enumerate}
    \item Applying the test, the court found that (1) the rule in question was 
    congressional, (2) it unavoidably conflicted with the Massachusetts rule, 
    (3) it was constitutional, and (4) it was consistent with the Rules 
    Enabling Act. Thus, the federal rule applied.
    \item Reversed.
\end{enumerate}

\subsection{Applying the \emph{Hanna} Test: \emph{Walker v. Armco Steel Corp.}}

Under the \emph{Hanna} test, a federal procedural rule displaces a state rule 
only if there is an unavoidable conflict between the two.

\begin{enumerate}
    \item A nail head fragmented and injured Walker's eye. He sued the 
    manufacturer, Armco, in Oklahoma district court.
    \item The district court dismissed the action on the grounds that under 
    Oklahoma procedural rules, the statute of limitations (two years) and the 
    service period (60 days) had run. The Tenth Circuit affirmed.
    \item On appeal, Walker argued that the court should follow FRCP 3, which 
    states only that a civil action commences when the plaintiff files a 
    complaint with a court.
    \item Justice Marshall:
    \begin{enumerate}
        \item Walker argued that there is a direct conflict between the state 
        and federal procedural rules. Under \emph{Hanna}, this would mean that 
        the federal rule prevails.
        \item In this case, however, there was no unavoidable conflict between 
        the state and federal rules. ``There is no indication that the 
        [federal] Rule was intended to toll a state statute of limitations,'' 
        much less displace state rules.\footnote{446 U.S. 740 (1980).}
        \item State law should have applied. Under the Oklahoma rule, the 
        statute of limitations had run. Affirmed.
    \end{enumerate}
\end{enumerate}

\subsection{Standards of Appellate Review: \emph{Gasperini v. Center for 
Humanities}}

State statutes regarding the standard of review for excessive jury verdicts 
are substantive, not procedural, and federal courts are therefore obligated to 
follow them.

\begin{enumerate}
    \item The Center for Humanities lost 300 of Gasperini's slides. He sued in 
    New York district court. The trial court awarded \$450,000 in compensatory 
    damages.
    \item A New York statute required reversal of an ``excessive'' jury award 
    if it ``deviates materially from what would be reasonable compensation.''
    \item Justice Ginsburg: the New York statute was a ``manifestly 
    substantive'' policy. The federal court must apply the state standard.  
    Remanded for a new trial so that the judge could apply the state standard 
    to the jury's verdict.
\end{enumerate}

\subsection{Federal Common Law: \emph{Clearfield Trust Co v. United States}}

Although there is no general federal common law, courts can still apply 
federal common law in particular circumstances.

\begin{enumerate}
    \item A paycheck to Clair Barner from the WPA was stolen, Barner's 
    signature forged, and cashed by Clearfield Trust at a JC Penney.
    \item The government sued Clearfield Trust in Pennsylvania district court. 
    The court determined that Pennsylvania state law controlled, and therefore 
    dismissed the complaint because the government was too slow to notify 
    Clearfield of the forgery.
    \item The Third Circuit reversed.
    \item Justice Douglas: the government's interest in uniformity and 
    certainty in its financial dealings weighed in favor of applying federal 
    law over state law. ``The issuance of commercial paper by the United 
    States is on a vast scale and transactions in that paper from issuance to 
    payment will commonly occur in several states. The application of state 
    law, even without the conflict of laws rules of the forum, would subject 
    the rights and duties of the United States to exceptional uncertainty. It 
    would lead to great diversity in results by making identical transactions 
    subject to the vagaries of the laws of the several states. The 
    desirability of a uniform rule is plain.'' Affirmed.
\end{enumerate}
