\section{Jury Trials}

\subsection{Overview and Background}

\begin{enumerate}
    \item Very few cases go to trial.\footnote{Casebook p. 1057.}
    \item As of 2002, 2/3rds of federal trials were heard before a jury, while 97\% of all state trials were heard before a judge.\footnote{Casebook p. 1058.}
    \item Pre--jury verdict motions: nonsuit, directed verdict (state) or judgment as a matter of law (federal).
    \item Post--jury verdict motions: judgment n.o.v. (state), judgment as a matter of law (federal), or new trial (both).\footnote{Casebook p. 1060.}
    \item A party waives the right to a jury trial if it does not request it. FRCP 38.
    \item Seventh Amendmend guarantee to a jury trial does not apply to the states.\footnote{Casebook p. 1098.}
    \item Six and twelve are the most common jury sizes. Six is the lower limit.
    \item The Supreme Court has upheld non-unanimous verdicts. FRCP 48 requires a unanimous verdict unless the parties stipulate otherwise.
\end{enumerate}

\subsection{Judgment as a Matter of Law: \emph{Simblest v. Maynard}}

\begin{enumerate}
    \item A travelling salesman collided with a fire truck in an intersection in Burlington. He claimed that the power went out when he was in the middle of the intersection, including the traffic lights. All other witnesses testified that the power had gone out at least ten minutes before.
    \item He looked to his right when he was half or three-fourths of the way through the intersection. That was the first time he saw the fire truck. He claimed to have not heard the siren nor seen the flashing lights.
    \item The salesman sued for negligence to recover for injuries he suffered from the collision.
    \item At the close of the plaintiff's case, the defendant moved for a directed verdict. Denied.
    \item The jury found for the plaintiff. The defendant moved for judgment n.o.v., which the judge granted.
    \item The plaintiff argued, first, that the district court granted the judgment n.o.v. in error.
    \begin{enumerate}
        \item The court here explained the standard of review for judgments n.o.v. \textbf{``that, without weighing the credibility of the witnesses or otherwise considering the weight of the evidence, there can be but one conclusion as to the verdict that reasonable men could have reached.''}\footnote{Casebook p. 1165.} Evidence must be viewed in the light most favorable to the non-moving party.
        \item A key question is whether the court con consider \emph{all} evidence or only evidence favorable to the non-moving party and all uncontradicted, unimpeached evidence against him. Vermont law would allow all evidence. The plaintiff argued that the federal rules would allow only favorable evidence. The court here held that the federal standard also allows admission of uncontradicted, unimpeached evidence.
        \item Under either rule, however, the court believed it was clear that the plaintiff was guilty of contributory negligence. Thus, the judgment n.o.v. for the defendant was properly granted.
    \end{enumerate}
    \item The plaintiff argued in the alternative that the district court erred in declining to instruct the jury on last clear chance. The court dispatched with this argument with the conclusion that ``the overwhelming, uncontroverted evidence demonstrates that defendant in the exercise of due care simply could not have avoided the accident.''\footnote{Casebook 1169.}
    \item The \emph{Simblest} court developed a three-part test for determining whether a jury could reasonably find for the non-moving party:\footnote{Casebook p. 1170 n. 2.}
    \begin{enumerate}
        \item Review all of the evidence in the record. [---but didn't the court interpret the federal standard to only include unfavorable evidence if it's unimpeached and uncontradicted?]
        \item Draw all reasonable inferences in favor of the non moving party.
        \item Not make credibility determinations or weigh the evidence.
    \end{enumerate}
    \item Courts rarely grant pre-verdict motions for judgment as a matter of law, because if the case is successfully appealed, it will have to be retried entirely from the beginning. However, if the judge grants a motion for judgment after a matter of law after the jury has returned a verdict, a successful appeal will only reinstate that jury's verdict.
\end{enumerate}

\subsection{\emph{Sioux City \& Pac R.R. Co. v. Stout}}

\begin{enumerate}
    \item A six-year-old was injured while playing on an unlocked railyard turntable. His parents brought an action for negligence. The jury returned a verdict of \$7,500 for the plaintiff.
    \item The court reasoned that if the evidence led to a reasonable inference of negligence, it was free to find for the plaintiff. In this case, because it was reasonable to think that an unlocked turntable might lead to injuries for kids playing on it, negligence was a reasonable inference.
    \item ``...although the facts are undisputed it is for the jury and not for the judge to determine whether proper care was given, and whether they establish negligence.''\footnote{Casebook p. 1175.}
    \item What distinguishes this case from \emph{Simblest}? The facts are disputed in both. In both, the court held in substance that the facts were in dispute. In \emph{Simblest}, however, the judge did not believe that the facts could give rise to a reasonable finding in favor of the plaintiff. In this case, however, the judge believed it was reasonable to infer the defendant's negligence.
\end{enumerate}

\subsection{FRCP 50: Judgment as a Matter of Law in a Jury Trial; Related Motion for a New Trial; Conditional Ruling}

\begin{itemize}
    \item (a) Judgment as as matter of law.
    \item (b) Renewing the motion after trial; alternative motion for a new trial.
    \item (c) Granting the renewed motion; conditional ruling on a motion for a new trial.
    \item (d) Time for a losing party's new-trial motion.
    \item (e) Denying the motion for judgment as a matter of law; reversal on appeal.
\end{itemize}

\subsection{FRCP 59: New Trial; Altering or Amending a Judgment}

\begin{enumerate}
    \item (a) In general.
    \item (b) Time to file a motion for a new trial.
    \item (c) Time to serve affidavits.
    \item (d) New trial on the court's initiative or for reasons not in the motion.
    \item (e) Motion to alter or amend a judgment.
\end{enumerate}

\subsection{Overview of Post-Trial Motions}

\begin{enumerate}
    \item A court can do two things after a jury returns a verdict:\footnote{Casebook p. 1180.}
    \begin{enumerate}
        \item Enter judgment against the verdict-winner---i.e., judgment as a matter of law (federal law) or judgment n.o.v. (state).
        \item Grant a new trial because of weak evidence or procedural errors.
    \end{enumerate}
\end{enumerate}

\subsection{\emph{Tanner v. United States}}

\begin{enumerate}
    \item Petitioners Conover and Tanner were convicted of conspiring to defraud the United States. The Eleventh Circuit affirmed.
    \item The petitioners argued that the district court erred in refusing to admit juror testimony at a post-verdict hearing on juror intoxication during the trial.
    \item Evidence shows that multiple jurors were drinking, smoking marijuana, and taking cocaine throughout the trial. One of the jurors involved reported the activity to the petitioner's attorney. The district court denied motions for leave to interview jurors or for an evidentiary hearing.
    \item Justice O'Connor:
    \begin{enumerate}
        \item The firm common law rule is that juror testimony cannot impeach a jury verdict. A few exceptions exist for ``external'' influence, e.g., when a juror had applied for a job in the D.A.'s office. Juror testimony is not allowed on ``internal'' influences, including allegations of insanity or lack of understanding of English.
        \item The rule exists to protect jurors from harassment and to preserve the jury's independence.
        \item ``There is little doubt that post-verdict investigation into juror misconduct would in some instances lead to the invalidation of verdicts reached after irresponsible or improper juror behavior. It is not at all clear, however, that the jury system could survive such efforts to perfect it.''\footnote{Casebook p. 1184.}
        \item The legislative history of the Federal Rules of Evidence suggest that Congress explicitly rejected a version of the rules ``that would have allowed jurors to testify on juror conduct during deliberations, including juror intoxication.''
        \item Petitioners argue that the refusal to hold an evidentiary hearing at which jurors could testify violates the Sixth Amendment guarantee to a competent jury. The court held that nonjuror evidence and the inadmissibility of juror testimony rendered a post-verdict evidentiary hearing unnecessary.
    \end{enumerate}
    \item Justice Marshall, dissenting:
    \begin{enumerate}
        \item The Federal Rules of Evidence (here, 606(b)) does not preclude juror testimony on conduct before or after deliberations. Because the allegations of misconduct here involved juror activity beyond deliberation, it should have been admitted.
        \item Justice Marshall has a different reading of the legislative history.
        \item Regarding the internal/external influence distinction, as ``a common sense matter, drugs and alcohol \emph{are} outside influences on jury members.
    \end{enumerate}
\end{enumerate}

\subsection{\emph{Spurlin v. Gen. Motors}}

\begin{enumerate}
    \item 
\end{enumerate}

\subsection{Entering and Enforcing Judgments}

\begin{enumerate}
    \item 
\end{enumerate}


