\section{Pleading}

\begin{enumerate}
    \item FRCP 8 guides the \textbf{substance} of pleadings, requiring (a) a 
    plaintiff's pleading to contain a statement of jurisdiction, a statement 
    of the claim, a demand for relief, and (b) a response to include a defense 
    to each claim, admission or denial of each allegation (or a general 
    denial).
    \item FRCP 10 guides the \textbf{form} of pleadings.
    \item In \emph{Swierkiewicz}, the court interpreted the pleading 
    requirements under FRCP 8 to be loose and liberal. The pleader need only 
    state a legally cognizable claim. There is no requirement to produce 
    evidence.
    \item Before \emph{Twombly}, a pleading was sufficient if it was 
    \textbf{conceivable}. Now, a claim must be \textbf{plausible}. The full 
    \emph{Twombly} standard:
    \begin{enumerate}
        \item Pleading must contain a short and plain statement (Rule 8).
        \item The court accepts facts as true.
        \item The court draws all reasonable inferences in favor of the 
        plaintiff.
        \item The court determines whether the complain is \emph{plausible on 
        its face}, meaning it must be more than a mere possibility and not 
        merely consistent with the defendant's liability.
    \end{enumerate}
    \item \textbf{Pleading in the alternative} allows multiple inconsistent 
    claims. \emph{McCormick}.
    \item Misconduct is sanctionable under \textbf{Rule 11}. \emph{Zuk}.
    \item FRCP 8(b) requires responses to claims to explicitly affirm or deny 
    each element of a claim. \emph{Zielinski}.
    \item Complaints can be \textbf{amended} before and during the trial under 
    the conditions of Rule 15.
\end{enumerate}

\subsection{FRCP 8: General Rules of Pleading}

\begin{itemize}
    \item (a) Pleading must contain:
    \begin{itemize}
        \item (1) Short and plain statement of jurisdiction.
        \item (2) Short and plain statement of claim.
        \item (3) Demand for relief.
    \end{itemize}
    \item (b) Defenses; Admissions and Denials:
    \begin{itemize}
        \item (1) General response:
        \begin{itemize}
            \item (A) Short and plain statement of defense to each claim.
            \item (B) Admit or deny allegations.
        \end{itemize}
        \item (2) Denial must respond to substance of allegation.
        \item (3) General denial will deny everything. Otherwise, specific 
        allegations must be separately admitted or denied.
        \item (4) Denial in part requires admission of the part that's true 
        and denial of the rest.
        \item (5) Defense must acknowledge lack of knowledge where it exists, 
        which has the effect of a denial.
        \item (6) Failure to deny = admission.
    \end{itemize}
\end{itemize}

\subsection{FRCP 10: Form of Pleadings}

\begin{itemize}
    \item (a) Caption/names of parties.
    \item (b) Paragraphs must be numbered; independent statements must be 
    separated.
    \item (c) Adoption by reference.
\end{itemize}

\subsection{Liberal Pleading Requirements: \emph{Swierkiewicz v. Sorema, N.A.}}

Complaints need not contain evidentiary details. They \enquote{must contain 
only \enquote{a short and plain statement of the claim showing that the 
pleader is entitled to relief.}}

\begin{enumerate}
    \item Swiekiewicz alleged employment discrimination based on his age and 
    nationality.
    \item The Southern District of New York dismissed for having ``not alleged 
    a prima facie case, in that he had not alleged circumstances supporting an 
    inference of discrimination.''\footnote{Casebook p. 609.} The Second 
    Circuit affirmed.
    \item Justice Thomas:
    \begin{enumerate}
        \item The lower courts' reliance on \emph{McDonnell Douglas} was 
        misplaced because that case established an evidentiary requirement, 
        not a \emph{pleading} requirement.
        \item Rule 8's simple standard should apply here. In \emph{Conley}, 
        the Court held that ``The Federal Rules reject the approach that 
        pleading is a game of skill.''\footnote{Casebook p. 612.}
        \item Reversed.
    \end{enumerate}
\end{enumerate}

\subsection{Plausible Pleadings: \emph{Bell Atlantic v. Twombly}}

Plaintiffs must state a plausible claim, not merely a conceivable one.

\begin{enumerate}
    \item The plaintiffs alleged that ILECs inflated charges for local phone 
    and high-speed Internet service by (1) engaging in ``parallel conduct'' to 
    inhibit startup CLECs and (2) agreeing not to compete with each other. The 
    plaintiffs provided little factual basis for their claims.
    \item The Southern District of New York dismissed for failure to state a 
    claim. The Second Circuit reversed.
    \item Justice Souter:
    \begin{enumerate}
        \item In complaints, ``[f]actual allegations must be enough to raise a 
        right to relief \emph{above the speculative 
        level}.'' The complaint needs ``enough fact 
        to raise a reasonable expectation that discovery will reveal evidence 
        of illegal agreement.''\footnote{Casebook p. 620.}
        \item Discovery can be expensive, and in this case, especially 
        expensive.
        \item Reversed.
    \end{enumerate}
\end{enumerate}

\subsection{Affirming \emph{Twombly}: \emph{Ashcroft v. Iqbal}}

\begin{enumerate}
    \item Iqbal was detained after 9/11. His complaint against Ashcroft and 
    Mueller alleged unconstitutionally harsh confinement on account of his 
    race, religion, or national origin.
    \item The district court denied Ashcroft's motion to dismiss for failure 
    to state a claim. The Second Circuit affirmed.
    \item Justice Kennedy:
    \begin{enumerate}
        \item After \emph{Twombly}, Rule 8(a)(2) requires ``further factual 
        enhancement.''
        \item The claim must be plausible. Iqbal's was not. Reversed.
    \end{enumerate}
\end{enumerate}

\subsection{Pleading in the Alternative: \emph{McCormick v. Kopmann}}

Parties can plead claims that are fundamentally inconsistent. Pleadings are 
not binding admissions. They're hypotheses subject to proof and disproof.

\begin{enumerate}
    \item McCormick, the widow of a man killed in a collision with Kopmann, 
    made two claims that were at issue on appeal:
    \begin{enumerate}
        \item Count I: wrongful death against Kopmann.
        \item Count IV: dram shop claim against the Huls.
    \end{enumerate}
    \item The defendants moved for directed verdicts. Denied. The jury 
    returned a verdict for \$15,000 against Kopmann and found the Huls not 
    guilty. Kopmann moved for judgment notwithstanding the verdict or for a 
    new trial; both were denied.
    \item Kopmann appealed, arguing:
    \begin{enumerate}
        \item Counts I and IV were mutually exclusive. The court agreed, but 
        held that this does not prevent them from being pleaded together. The 
        Illinois Civil Practice Act (modeled on FRCP 8(e)(2), which allows 
        alternative pleading where the plaintiff is ``genuinely in doubt as to 
        what the facts are and what the evidence will 
        show''\footnote{Caseboko p. 663.}).
        \item Allegations of intoxication in Count IV count as binding 
        admissions. Court held that ``he is not `admitting' anything other 
        than his uncertainty.''\footnote{Casebook p. 664.}
        \item Prima facie case in Count IV means plaintiff was contributorily 
        negligent regarding count one. The court found that the plaintiff did 
        exercise due care and was not intoxicated.
    \end{enumerate}
    \item Affirmed.
\end{enumerate}

\subsection{Sanctions: \emph{Zuk v. E. Penn. Psychiatric Institute}}

Misconduct is sanctionable under Rule 11.

\begin{enumerate}
    \item Zuk sued EPPI for copyright infringement. The district court dismissed 
    under FRCP 12(b)(6) (motion to dismiss) and found Zuk and his counsel 
    liable for \$15,000 in sanctions and counsel fees. Zuk settled his 
    liability. His counsel, Lipman, appealed.
    \item EPPI moved for sanctions ``on the grounds essentially that appellant 
    had failed to conduct an inquiry into the facts reasonable under the 
    circumstances and into the law.''\footnote{Casebook p. 671.} The court 
    imposed sanctions based on FRCP 11 and 28 U.S.C.\S\ 1927.
    \item The Third Circuit held:
    \begin{enumerate}
        \item Lipman had no liability under the Copyright Act.
        \item There was no bad faith on Lipman's part. Therefore, the lower 
        court abused its discretion in awarding sanctions based on 28 U.S.C. 
        \S\ 1927.
        \item The lower court did not identify exactly how the sanctions were 
        based on the FRCP and USC rules. The FRCP sanctions were appropriate, 
        because the original plaintiffs' research was deficient both factually 
        and legally, but it is not possible to review them further. Remanded 
        to consider the type and amount of sanctions specific to the FRCP 
        violation.
        \item ``~.~.~.~we conclude that it was in error to invoke without 
        comment a very severe penalty.''\footnote{Casebook p. 678.}
    \end{enumerate}
\end{enumerate}

\subsection{FRCP 11: Signing, Representations, Sanctions}

\begin{itemize}
    \item (a) Attorney (or self-represented party) must sign all papers.
    \item (b) Representations must:
    \begin{itemize}
        \item (1) Not have any improper purpose.
        \item (2) Make claims warranted by law or make nonfrivolous arguments 
        for changing the law.
        \item (3) Have evidentiary support (or probably support) for factual 
        contentions.
        \item (4) Make denials of factual contentions warranted by evidence, 
        reasonable belief, or lack of information.
    \end{itemize}
    \item (c) Sanctions:
    \begin{itemize}
        \item (1) After notice and opportunity to be heard, court may sanction 
        an attorney, firm, or party for 11(b) violations. Law firms must be 
        held jointly responsible.
        \item (2) Motions for sanctions must be made separately. The party has 
        21 days from the date of service to withdraw the challenged paper. The 
        court can award expenses to the prevailing party.
        \item (3) On its own initiative, a court can ask a party to show why 
        it hasn't violated 11(b).
        \item (4) Sanctions must be limited to what is necessary for 
        deterrence. Sanctions can be nonmonetary or monetary.
        \item (5) Limits on monetary sanctions:
        \begin{itemize}
            \item Courts cannot sanction a represented party for violating 
            11(b)(2) (i.e., lawyers are responsible for legal contentions).
            \item Courts cannot impose monetary sanctions unless it issues an 
            11(c)(3) show-cause order before dismissal or settlement.
        \end{itemize}
        \item (6) A sanction order must include an explanation.
    \end{itemize}
    \item (d) Rule 11 does not apply to discovery motions.
\end{itemize}

\subsection{FRCP 12: Defenses and Objections}

\subsection{Explicit Affirmation or Denial: \emph{Zielinski v. Philadelphia 
Piers}}

FRCP 8(b) requires responses to claims to explicitly affirm or deny each 
element of a claim.

\begin{enumerate}
    \item Facts:
    \begin{enumerate}
        \item February 9, 1953: Sandy Johnson crashed a forklift into Frank 
        Zielinski, causing injuries.
        \item February 10, 1953: Carload Contractors, Inc. sent an accident 
        report to its insurance company.
        \item April 28, 1953: Zielinski filed complaint against Philadelphia 
        Piers, Inc., arguing that (1) PPI owned the forklift that Johnson was 
        operating and (2) Johnson was acting as an employee of PPI.
        \item April 29, 1953: complaint was forwarded to the insurance 
        company.
        \item June 12, 1953: PPI's general manager responded to 
        interrogatories 1 through 5.
        \item August 19, 1953: Sandy Johnson testified in a deposition that he 
        was an employee of PPI.
        \item September 27, 1955: Pre-trial conference held where Zielinski 
        learned that the work of moving freight on the piers had been sold to 
        Carload Contractors, Inc., and that Johnson became an employee of 
        Carload.
        \item October 21, 1955: From the answers to supplementary 
        interrogatories, plaintiff learned that the accident report had been 
        submitted to the insurance company.
    \end{enumerate}
    \item The district court held:
    \begin{enumerate}
        \item In the original complaint, Zielinski claimed that PPI owned the 
        forklift and that Johnson was an employee of PPI. Initially, PPI 
        simply responded, ``Defendant~.~.~.~denies the averments of paragraph 
        5.''\footnote{Casebook p. 686.}
        \item PPI's response was inadequate. It meant to only deny the claim 
        of employee negligence, but it should have also addressed the question 
        of whether PPI owned and operated the forklift. PPI's response led 
        Zielinski to believe that PPI was the owner and operator of the 
        forklift, but in fact it had been sold to CCI. Zielinski sued the 
        wrong company, but by the time he learned his mistake (September 27, 
        1955), the statute of limitations had run.
        \item FRCP 8(b) requires responses to claims to explicitly affirm or 
        deny each element of a claim.
        \item The court held that the doctrine of equitable estoppel requires 
        that the case go forward against PPI with the record showing that PPI 
        owned and operated the forklift---even though that wasn't actually the 
        case.
    \end{enumerate}
\end{enumerate}

\subsection{FRCP 15: Amended and Supplemental Pleadings}

\begin{itemize}
    \item (a) Amendments before trial.
    \item (b) Amendments during and after trial.
    \item (c) Relation back of amendments.
    \item (d) Supplemental pleadings.
\end{itemize}

\subsection{Amendments Relating Back: \emph{Worthington v. Wilson}}

Complaints can be amended to name a different party as long as the new party 
had notice.

\begin{enumerate}
    \item Facts:
    \begin{enumerate}
        \item February 25, 1989: Worthington was arrested by two police 
        officers.
        \item February 25, 1991: Worthington filed a complaint against the 
        Village of Peoria Heights in county court against ``three unknown 
        named police officers.'' The Village removed the action to the 
        District Court for the Central District of Illinois.
        \item June 17, 1991: Worthington amended his complaint to identify the 
        two police officers by name.
    \end{enumerate}
    \item The statute of limitations was two years, which had expired when 
    Worthington filed his amended complaint. Defendants argued that the 
    amended complaint cannot related back to the filing date of the original 
    complaint under FRCP 15(c).
    \item In response, Worthington made three arguments:
    \begin{enumerate}
        \item Relation back was governed by an Illinois statute, not 15(c).
        \item The 15(c) requirements were satisfied.
        \item He should not be punished for omitting the officers' names in 
        the original complaint because the Peoria Heights Police Department 
        had withheld that information.
    \end{enumerate}
    \item In \emph{Schiavone v. Fortune} (1986), the Supreme Court held that 
    parties to be brought in by amendment must receive notice of the action 
    before the expiration of the statue of limitations period.
    \item December 1, 1991: FRCP 15 was amended in direct response to 
    \emph{Schiavone}. The new rule allowed complaints to be amended at any 
    time to correct misnamed defendants as long as the defendant was aware of 
    the action within 120 days of the filing of the original complaint. Notice 
    is no longer required.
    \item Under the new version of FRCP 15(c), Worthington's amendment was 
    acceptable.
    \item Defendants also argued that 15(c) did not apply because the defect 
    in Worthington's original complaint was due to lack of knowledge, not a 
    ``mistake.'' The Seventh Circuit had previously held that lack of 
    knowledge of the proper defendant does not involve a mistake. The district 
    court here disagreed, but out of deference agreed with defendants and 
    granted their motion to dismiss.
\end{enumerate}
