\section{Provisional Remedies}

\subsection{Statute of Limitations: \emph{United States v. Kubrick}}

\begin{enumerate}
    \item When does the statute of limitations clock begin?
    \item Plaintiff was rendered partially deaf from neomycin treatment at the 
    VA. He discovered the possibility of malpractice only after the two year 
    statute of limitations had expired.
    \item The issue was whether the claim accrues when the plaintiff is aware 
    of the existence and cause of his injury or when he is also aware of the 
    possibility of malpractice. The Supreme Court (White) holds the former 
    (Stevens dissenting).
    \item Rule vs. standard:
    \begin{enumerate}
        \item White: there's a clear rule here.
        \item Stevens: a rule is unnecessary---all we need is a looser 
        standard that can be applied on a case-by-case basis.
    \end{enumerate}
    \item Why does the statute of limitations exist? (see CB 54-55)
    \begin{enumerate}
        \item Protect against the ``cloud of litigation''
        \item Protect against ``stale claims''
        \item Keep the plaintiff from sitting on his rights
    \end{enumerate}
\end{enumerate}

Themes throughout the class:

\begin{enumerate}
    \item Procedure as \emph{policy}: how does procedure express values about 
    justice?
    \item Procedure as \emph{strategy}: how do actors use process 
    strategically?
    \item Procedure and \emph{power}: whom do rules benefit? Does access to 
    rulemakers matter? Why have rules at all? How do state vs. federal issues 
    come into play?
\end{enumerate}

\subsection{Due Process Requirements}

\begin{enumerate}
    \item Fifth Amendment: ``No person shall be ... deprived of life, liberty, 
    or property, without due process of law''
    \item Fourteenth Amendment: ``No \textbf{State} shall ... deprive any 
    person of life, liberty, or property, without due process of law''
    \item Remedies:
    \begin{enumerate}
        \item \emph{Plenary}: Usually awarded at the end of a lawsuit. Usual 
        types: compensatory and punitive damages, injunctions, and declaratory 
        judgments.
        \item \emph{Provisional}: Can be awarded at any time while a lawsuit 
        is pending. Usual types: attachment (seizure of property), temporary 
        restraining orders, preliminary injunctions. They are ``designed to 
        stabilize the situation pending the final disposition of the case or 
        to provide security to the plaintiff so that if she succeeds in 
        obtaining judgment she will be able to enforce it 
        effectively.''\footnote{Casebook p. 46.}
    \end{enumerate}
\end{enumerate}

\subsection{Writs of Replevin: \emph{Fuentes v. Shevin}}

\begin{enumerate}
    \item What are the due process requirements for ex parte decisions?
    \item Do statutes that allow writs of replevin only upon ex parte 
    application and posting of bond violate the Fourteenth Amendment?
    \item \emph{Fuentes v. Shevin} In multiple consolidated cases, a creditor 
    was granted, ex parte, a writ of replevin for the property of a debtor in 
    default (which was allowed by statutes in Florida and Pennsylvania). The 
    court found that these statutes violate the Due Process Clause (Fourteenth 
    Amendment). Absent extraordinary circumstances, due process requires 
    notice and hearing before deprivation.
    \begin{enumerate}
        \item Justice White: Parties whose rights are to be affected are 
        entitled to be heard; and in order that they may enjoy that right they 
        must must first be notified. The right to be heard is a basic part of 
        procedural fairness.
    \end{enumerate}
    \item Key due process protections: \textbf{notice} and \textbf{opportunity 
    to be heard} (in a meaningful way)
\end{enumerate}

Similar cases:

\begin{enumerate}
    \item \emph{Mitchell v. W.T. Grant}: A similar statute in Louisiana was 
    upheld on the grounds that (1) the applicant for the writ must declare the 
    facts in a certified petition or affidavit, and (2) the showing must be 
    made to a judge, not merely a court official.
    \item \emph{North Georgia Finishing, Inc. v. Di-Chem, Inc.}: A similar 
    Georgia statute was struck down because (1) the affidavit can be filed by 
    the petitioner's attorney, who need not have any direct knowledge of the 
    facts of the dispute, and (2) the writ is issuable by a court clerk, not a 
    judge.
\end{enumerate}

The minimum constitutional requirements for valid ex parte prejudgment and 
seizure appear to be:

\begin{enumerate}
    \item An application grounded in facts.
    \item Issued by a judge, not a court official.
    \item A speedy hearing.
    \item Only applicable to a limited range of transactions\footnote{See 
    California's version of these statues, Casebook p. 82 top.}.
\end{enumerate}

\subsection{Prejudgment Attachment of Real Estate: \emph{Doehr}}

\begin{enumerate}
    \item Does prejudgment attachment of real estate (a) without prior notice 
    or hearing, (b) without extraordinary circumstances, and (c) without a 
    bond satisfy the Due Process Clause in the Fourteenth Amendment? 
    \emph{Doehr}: no.
    \item \emph{Connecticut v. Doehr}: Under a CT statute, DiGiovanni won a 
    \$75,000 prejudgment attachment on Doehr’s home in conjunction with a 
    civil action for assault and battery. The court unanimously held that the 
    CT statute would allow deprivation for cases where the defendant's 
    property claim would fail to convince a jury. Without exigent 
    circumstances, a preattachment hearing is required.
    \item Concurrences:
    \begin{enumerate}
        \item Marshall, Stevens, O'Connor, White: Bonds are also necessary in 
        all cases.
        \item Rehnquist, Blackmun: Liens can serve a useful purpose (e.g., for 
        laborers to enforce their interests over delinquent landowners). Also, 
        the terms ``bond'' and ``exigent circumstances'' are overly vague.
    \end{enumerate}
    \item \emph{Matthews} rules provide a test for whether deprivation meets 
    due process requirements:
    \begin{enumerate}
        \item Consideration of the private interests that the deprivation will 
        affect.
        \item Examination of the risk of erroneous deprivation, and the 
        safeguards in place.
        \item Attention to the interest of the party seeking the judgment 
        remedy (in \emph{Mathews} originally, this was the government; here 
        it's the private plaintiff).\footnote{Casebook p. 85.}
    \end{enumerate}
\end{enumerate}
